Synthesis report

### 2020-05-17 analysis/reports/figs.Rmd
we may have a draft?????

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log150_nparts = log(gmp:::as.double.bigz(nparts), base = 150)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found==1) %>%
  filter(s0 >= 2,
         n0 > s0,
         dat %in% c("bbs", "gentry", "fia_short", "fia_small", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(substr(dat, 0, 3) == "fia", yes = "fia", no = dat))
```

@@

```{r datasets in s and n space}
ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_d(end = .9)

@@

@@

# Binned by ranked nparts and ranked 95 intervals

```{r binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

@@

```{r binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_2t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

@@

```{r relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_2t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_2t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

# 95 interval vs nparts

```{r 1t 95 interval}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log(s0), y = log(n0), color = skew_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

ggplot(filter(all_di, singletons == F), aes(x = log(s0), y = log(n0), color = simpson_95_ratio_1t)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_c(option = "magma", end = .9, begin = .1)

```{r 1t 95 ratio v nparts scatter}

ggplot(filter(all_di, singletons == F, s0 > 2), aes(x = log_nparts, y = skew_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

ggplot(filter(all_di, singletons == F), aes(x = log_nparts, y = simpson_95_ratio_1t, color = log(n0/s0))) +
  geom_point() +
  theme_bw() +
  geom_vline(xintercept = c(10,15)) +
  scale_color_viridis_c()

```{r 1t percentiles v 95 interval}

ggplot(filter(all_di, !singletons), aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw()

# Binned by ranked nparts and ranked 95 intervals

```{r 1t binned skew}

all_di_skew <- all_di %>%
  filter(!singletons, s0 > 2) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(skew_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: Nparts")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("1/10th bins: 95% interval ratio")
ggplot(all_di_skew, aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r 1t binned simpson}

all_di_simpson <- all_di %>%
  filter(!singletons) %>%
  arrange(log_nparts) %>%
  mutate(nparts_rank = row_number()) %>%
  arrange(simpson_95_ratio_1t)%>%
  mutate(ratio_rank = row_number(),
         random_rank = sample.int(n = length(nparts_rank), size = length(nparts_rank), replace = F)) %>%
  mutate(nparts_rank_binned = ceiling(10 * (nparts_rank / max(nparts_rank))),
         ratio_rank_binned = ceiling(10 * (ratio_rank / max(ratio_rank))),
         random_rank_binned = ceiling(10 * (random_rank / max(random_rank))))

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(nparts_rank_binned), scales = "free") +
  theme_bw() +
    ggtitle("1/10th bins: Nparts")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(ratio_rank_binned), scales = "free") +
  theme_bw()+
  ggtitle("1/10th bins: 95% interval ratio")

ggplot(all_di_simpson, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(random_rank_binned), scales = "free") +
  theme_bw() +
  ggtitle("10 random bins")

```{r 1t relationships in nparts bins}

ggplot(all_di_skew, aes(x = skew_95_ratio_1t, y = skew_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t, y = simpson_percentile)) +
  geom_point(alpha = .05) +
  theme_bw() +
  facet_wrap(vars(nparts_rank_binned), scales = "free")

```{r overall relationships}

ggplot(filter(all_di, !singletons), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw()

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()

### Remove FIA

```{r overall relationships no fia}

ggplot(filter(all_di, !singletons, dat != "fia_short", dat != "fia_small"), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw()

ggplot(filter(all_di, !singletons, dat != "fia_short", dat != "fia_small"), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()

```{r overall relationships by da t}

ggplot(filter(all_di, !singletons, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(vars(dat), scales = "free_y")

ggplot(filter(all_di, !singletons), aes(x = simpson_percentile)) +
  geom_histogram() +
  theme_bw()+
  facet_wrap(vars(dat), scales = "free_y")

```{r gentry}

ggplot(filter(all_di, dat == "gentry", !singletons), aes(x = log(n0/s0), y = simpson_percentile)) +
  geom_point()

gentry_cat <- filter(all_di, dat == "gentry", !singletons) %>%
  mutate(low_avgn = log(n0/s0) < 1.25)

ggplot(gentry_cat, aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(low_avgn), scales = "free_y")

```{r 1t hists}

ggplot(all_di_skew, aes(x = skew_95_ratio_1t))  +
geom_histogram() +   theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_skew, aes(x = skew_95_ratio_1t)) +
  geom_histogram(aes(x = skew_percentile / 100, y = ..count.. * 5), fill = "green", alpha = .5) + geom_histogram(alpha = .5) +  
theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t))  +
geom_histogram() +   theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

ggplot(all_di_simpson, aes(x = simpson_95_ratio_1t)) +
  geom_histogram(aes(x = simpson_percentile / 100, y = ..count.. * 5), fill = "green", alpha = .5) + geom_histogram(alpha = .5) +  
theme_bw() +
  facet_wrap(vars(nparts_rank_binned))

>>>>

all_di <- all_di %>%
  mutate(log_nparts = log(gmp:::as.double.bigz(nparts)),
         log_nsamples = log(nsamples),
         log150_nparts = log(gmp:::as.double.bigz(nparts), base = 150)) %>%
  mutate(prop_found = exp(log_nsamples - log_nparts)) %>%
  mutate(found_all = prop_found==1) %>%
  filter(s0 >= 2,
         n0 > s0,
         dat %in% c("bbs", "gentry", "fia_short", "fia_small", "mcdb", "misc_abund_short")) %>%
  mutate(dat = ifelse(substr(dat, 0, 3) == "fia", yes = "fia", no = dat)) %>%
  mutate(dat = ifelse(dat == "misc_abund_short", "misc_abund", dat))%>%
  mutate(is_fia = ifelse(substr(dat, 0, 3) == "fia", dat, "other"))
```

@@

```{r datasets in s and n space}

all_dat_to_plot <- filter(all_di, singletons == F) %>%
  bind_rows(mutate(filter(all_di, singletons == F), dat = "all"))

ggplot(all_dat_to_plot, aes(x = log(s0), y = log(n0), color = dat)) +
  geom_point(alpha = .5) +
  theme_bw() +
  theme(legend.position = "top") + scale_color_viridis_d(end = .9) +
  facet_wrap(vars(dat))

@@

skew_outcomes <- filter(all_di, !singletons, s0 > 2) %>%
  group_by(is_fia) %>%
  summarize(high_skew_prop = mean(skew_percentile >= 95)) %>%
  ungroup()

even_outcomes <- filter(all_di, !singletons) %>%
  group_by(is_fia) %>%
  summarize(low_even_prop = mean(simpson_percentile <= 5)) %>%
  ungroup()

@@

# Within the FIA size range

```{r within FIA size range}

fia_size_range <- filter(all_di, !singletons)

fia_max_s = max(filter(all_di, dat == "fia")$s0)
fia_max_n = max(filter(all_di, dat == "fia")$n0)

fia_size_range <- filter(fia_size_range,
                         s0 <= fia_max_s,
                         n0 <= fia_max_n)

skew_fia <- ggplot(filter(fia_size_range, s0 > 2), aes(x = skew_percentile)) +
  geom_histogram() +
  facet_wrap(vars(is_fia), scales = "free_y") +
  theme_bw() +
  ggtitle("Skew percentile - within FIA size range")

simpson_fia <- ggplot(filter(fia_size_range), aes(x = simpson_percentile)) +
  geom_histogram() +
  facet_wrap(vars(is_fia), scales = "free_y") +
  theme_bw() +
  ggtitle("Simpson percentile - within FIA size range")

gridExtra::grid.arrange(grobs = list(skew_fia, simpson_fia), nrow = 1)

@@

```{r 95 hists}

all_di <- all_di %>%
  group_by_all() %>%
  mutate(is_fia_size = (s0 <= fia_max_s) && (n0 <= fia_max_n)) %>%
  ungroup()

all_di <- all_di %>%
  mutate(is_fia_size = ifelse(is_fia_size, "FIA size", "Larger than FIA"))

skew_ratio <- ggplot(filter(all_di, s0 > 2), aes(x = skew_95_ratio_1t)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(vars(is_fia_size), scales = "free_y") +
  ggtitle("Narrowness of skewness")

even_ratio <- ggplot(filter(all_di, s0 > 2), aes(x = simpson_95_ratio_1t)) +
  geom_histogram() +
  theme_bw() +
  facet_wrap(vars(is_fia_size), scales = "free_y") +
  ggtitle("Narrowness of evenness")

gridExtra::grid.arrange(grobs = list(skew_ratio, even_ratio), nrow = 1)

@@

# Rarefaction

```{r rarefaction change}

rarefied <- filter(all_di, singletons) %>%
  select(dat, site, skew_percentile, simpson_percentile) %>%
  rename(skew_percentile_rarefied = skew_percentile,
         simpson_percentile_rarefied = simpson_percentile)

rarefied <- right_join(all_di, rarefied)

skew_rare <- ggplot(filter(rarefied, s0 > 2), aes(x = skew_percentile, y = skew_percentile_rarefied)) +
  geom_point(alpha = .2) +
  geom_abline(slope = 1, intercept = 0, color = "green") +
  theme_bw() +
  facet_wrap(vars(dat)) +
  ggtitle("Skew percentile change with rarefaction")

even_rare <- ggplot(filter(rarefied), aes(x = simpson_percentile, y = simpson_percentile_rarefied)) +
  geom_point(alpha = .2) +
  geom_abline(slope = 1, intercept = 0, color = "green") +
  theme_bw() +
  facet_wrap(vars(dat)) +
  ggtitle("Evenness percentile change with rarefaction")

gridExtra::grid.arrange(grobs = list(skew_rare, even_rare), nrow = 1)

skew_rare_outcome <- rarefied %>%
  filter(s0 > 2) %>%
  group_by(is_fia) %>%
  summarize(high_skew = mean(skew_percentile_rarefied >= 95)) %>%
  ungroup()
skew_rare_outcome

even_rare_outcome <- rarefied %>%
  filter(s0 > 2) %>%
  group_by(is_fia) %>%
  summarize(low_even = mean(simpson_percentile_rarefied <= 5)) %>%
  ungroup()
even_rare_outcome
```
